---
title: "Untitled"
date: "3/16/2022"
output:
html_document: default
pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning=FALSE}
library(dplyr)
library(spotifyr)
library(tidyverse)
library(ggplot2)
library(lubridate)
library(glmnet)
library(caret)
library(reshape2)
library(randomForest)
library(ModelMetrics)
```

```{r}
id <- '29dd370f3d934a20999183b22e985a7b'
secret <- "84fd28b593d84527bcafcfb0adab6106"
Sys.setenv(SPOTIFY_CLIENT_ID = id)
Sys.setenv(SPOTIFY_CLIENT_SECRET = secret)
access_token <- get_spotify_access_token()
```

```{r}
# data <- read.csv("clean_data//trackData.csv")
# #training_data <- data #%>% group_by_at(cols_to_keep)  #summarise(avg_rank = mean(ranking), weeks_trending = n()) %>% mutate(avg_rank_per_week = avg_rank/weeks_trending)# training data
dpath = "clean_data/trackData.csv"
data = read.csv(dpath) 
set.seed(1)
avg = data %>% 
group_by(song_name)  %>%  
summarise(s_freq = sum(ranking), weeks_trending = n()) %>% 
mutate(rankingX = s_freq / weeks_trending)
data = merge(avg, data, by="song_name")
data = select(data, -c(week, ranking, id,X,s_freq))
data=distinct(data)
```

```{r}
#This cell contains primarily contains functions that we use throughout the anaylsis process:


#////////////////////////////////////////////////////////////////////////////////////////////////

#Searchs spotify for a song and given a model predict is value using that model
#model_type
# elastic net/ lasso/ridge =1
# random forest = 2
# If it requires a lamda value then chose 1...

getEstimatedRank = function(song_query = "", mf = "", modelObj,  model_type, lambda = 0)
{
  new = spotifyr::search_spotify(song_query,type = "track")
  id = new["id"][[1]][1]
  track = get_track_audio_features(id)
  track = track[mf]
  new_track = matrix(as_vector(track),nrow = 1, ncol = length(mf))
    
  switch(model_type,
  "1"= 
  predict(modelObj,s = lambda, newx = new_track),
  "2"=
  predict(modelObj, newx = new_track, type = "response")
  )
}
#////////////////////////////////////////////////////////////////////////////////////////////////
#convert a set of values to 0,1  given a cutoff value ie 10 <= 50 therefore return 1
toBinary = function(data, cut_off) 
{
  binaryRep = sapply(data, function(rank){
    ifelse(rank < cut_off, 1, 0)
  })
  return(binaryRep)
}
#////////////////////////////////////////////////////////////////////////////////////////////////
#Give the lower triangle of a corrlation matrix
getLow = function(corMatrix) 
{
  corMatrix[upper.tri(corMatrix)] = NA
  return(corMatrix)
}
#Calculate and print mse, rmse, normalized rmse, and r squared value to console 
get_metrics = function(Actual, Predicted, model, mf, rf = 0){
  mse = mse(Actual,Predicted)
  rmse = rmse(Actual,Predicted)
  alt_rmse= sqrt(sum((Predicted+Actual)^2)/length(Actual))
  norm_rmse = rmse(Actual,Predicted) / (max(Actual)-min(Actual))
  tss = sum((Actual - mean(Actual))^2)
  rss = sum((Predicted-Actual)^2) #sum((Predicted - mean(Predicted))^2)
  if(rf==0)
  {
    r_sq = 1-(rss/tss)
  }
  else 
  {
    r_sq =mse/var(Actual)
  }
  top = (1-r_sq)*(length(Actual)-1)
  bottom = (length(Actual)-length(mf)-1)
  adj_r_sq = 1-(top/bottom)
  # sst = sum((Actual - mean(Actual))^2)
  # sse = sum((Predicted - Actual )^2)
  # rsq = 1 - sse/sst
  
  cat("\nGiven Model:",deparse(substitute(model)),"\n",
  "\n////////////////////////////////////////////",
  "\nMean squared error: ", mse,
  "\nRoot mean sqaured error: ",rmse,
  "\nNormailized root mean sqaured error: ",norm_rmse,
  "\nR^2 value of model: ",r_sq,
  "\nAdjusted R^2 value of model: ",adj_r_sq)
  cat(
  "\n////////////////////////////////////////////")
}
```

From here we want to take a look at the features that when compared to rank column have the strongest  correlations both negative and postive depending on the way in which the value is formatted. So if we look at the far right column we can see that very clearly there are postive correlations primarily between rank and danceability along with rank and valence. We generally want to look for negative correlations since the lower a songs rank is the closer it is to the number one spot on the charts. That being said within our data set it looks to be the case that energy, loudness, instrumentalness, livelness and duration are correlated with a "high" rank or is more likely to be on the upper section of the charts.


As disscussed above and shown by our correlation heatmap, the most relevant freatures to use to predict a song will be popular are energy, loudness, instrumentalness, livelness, and duration. Additionally to predict that song will be unpopular we also be adding in danceability, and valence to ensure the model is balanced. 

Now that brings us to the task of deciding on what kind machine learning model we should use in order to predict a songs popularity in Denver. Intailly we ended up trying lasso regression to automate the process of selecting features automated however this led to quite poor results. We also tried using ridge regression then elastic net but this also gave us less then perfect results. 

The code below is for that ridge/elastic net/lasso regression this and as can be seen in the console the the $R^2$ value is extremly low meaning that this certainly isn't the correct approach for this kind of a data set since it's quite noisey. 

```{r}
#selecting Features and making random traning/test partions of data 
#model_features = c(
# "energy",
# "loudness",
# "instrumentalness",
# "liveness",
# "duration_ms",
# "acousticness",
# "valence" )


model_features = c(
 "energy",
 "loudness",
 "instrumentalness",
 "liveness",
 "duration_ms",
 "acousticness",
 "valence",
 "key",
 "time_signature",
 "danceability",
 "tempo"
  )

train = data #
train$sample  = sample(c(1,0), replace=1, prob = c(0.7,0.2) , size = nrow(data)) # add sample column
test = train %>% filter( sample %in% c(0)) 
test = test[order(test$ranking),]
train = subset(train, sample!=0) 
true_rank = test$rank
```





The function below allows you to search for a song in the spotify catalog and using one of the models that we created output the prediction of that particular song. This particular call of the function is set to use the elastic net regression model.

```{r}
#still needs work
  # getEstimatedRank("Love Letter", model_features , random_forest_model, 2)
```

Logistical Regression 


```{r}
# View(train)
train_log <- train %>% select(is.numeric) 
train_log = train_log %>% mutate(bin_trend = toBinary(train_log$ranking, 50)) %>%  select(c(model_features,bin_trend))

test_log = test %>% select(is.numeric) 
test_log = test_log %>% mutate(bin_trend = toBinary(test_log$ranking, 50)) %>%  select(c(model_features,bin_trend))

log_reg_model <- glm(
  bin_trend~., 
  data = train_log, 
  family = "binomial"
)
pred <- predict(log_reg_model, test)
pred_binary = toBinary(pred,.5)
true_rank_binary = toBinary(true_rank, 50)

```

```{r}
log_reg_confusion = caret::confusionMatrix(as.factor(pred_binary), as.factor(true_rank_binary))
log_reg_confusion
```